\chapter{Methodology}
\label{ch:system-methodology}

\section{System Overview}
\label{sec:overview}

Our system takes open-source Python projects with broken tests as input and produces code patches that fix the bugs causing those failures. The pipeline has six stages: framework detection, hook injection, test execution with runtime tracing, filtering and serialization of the collected data, trace selection, and LLM-driven patch generation via a tool-based session. Figure~\ref{fig:pipeline} shows the full flow.

The starting point is a project instance in which at least one test fails. We evaluate on two benchmarks: SWE-bench Lite~\citep{jimenez2023swebench} and BugsInPy~\citep{widyasari2020bugsinpy}. In both cases the failing tests were introduced by actual commits that broke existing functionality, so the bugs are representative of real development work. When we evaluate on benchmarks, each test instance runs inside an isolated Docker container. A dedicated manager, \texttt{TracedInstanceRunner}, mounts three volumes into the container. The tracer scripts directory is mounted read-only at \texttt{/opt/tracers}, making the tracing code available inside the container without copying it into the image. A per-instance output directory is mounted read-write at \texttt{/trace\_output}; this is where \texttt{auto\_debug.json} is written after the tests finish, and the file is immediately accessible on the host once the run completes. A writable project snapshot is mounted at \texttt{/project\_mirror} and is used by the LLM tools to read source files. SWE-bench images pre-install the project source at \texttt{/testbed}; we copy that directory into \texttt{/project\_mirror} at container start, before any patch is applied, so the snapshot reflects the original buggy state. After the container is cleaned up, the host reads source files from this snapshot to build the source map that the LLM tools use when rendering code context.

\begin{figure}[ht]
\centering
\begin{tikzpicture}[
  node distance=0.55cm and 0cm,
  box/.style={
    rectangle, rounded corners=3pt,
    draw=black!70, fill=white,
    minimum width=6.8cm, minimum height=0.72cm,
    text centered, font=\small
  },
  arr/.style={-{Stealth[length=5pt]}, thick, draw=black!60}
]

\node[box] (A) {Project with failing tests};
\node[box, below=of A] (B) {Framework detection};
\node[box, below=of B] (C) {Hook injection into container};
\node[box, below=of C] (D) {Test execution (\texttt{sys.settrace})};
\node[box, below=of D] (E) {Runtime data collection};
\node[box, below=of E] (F) {Filtering pipelines};
\node[box, below=of F] (G) {Trace selection};
\node[box, below=of G] (H) {LLM tool session};
\node[box, below=of H] (I) {Patch submission};

\draw[arr] (A) -- (B);
\draw[arr] (B) -- (C);
\draw[arr] (C) -- (D);
\draw[arr] (D) -- (E);
\draw[arr] (E) -- (F);
\draw[arr] (F) -- (G);
\draw[arr] (G) -- (H);
\draw[arr] (H) -- (I);

\end{tikzpicture}
\caption{End-to-end pipeline from a failing test to a submitted patch.}
\label{fig:pipeline}
\end{figure}


\section{Framework Detection}
\label{sec:framework-detection}

Python projects do not use a single testing framework. The three most common runners are pytest, the standard-library unittest, and the Django test runner. Each framework invokes tests differently and exposes hooks through incompatible APIs. Injecting tracing without first identifying the runner would fail silently or produce empty output, so we classify the framework before any instrumentation takes place.

A \texttt{FrameworkDetector} inspects the project's evaluation script to assign one of four labels: \texttt{pytest}, \texttt{unittest}, \texttt{django}, or \texttt{unknown}. Detection works by scanning the eval script for known substrings: \texttt{"pytest"} or \texttt{"py.test"} indicate pytest, \texttt{"manage.py test"} or \texttt{"runtests.py"} indicate Django, and \texttt{"python -m unittest"} or \texttt{"unittest"} indicate the standard runner. If none of these match, the project is labelled unknown and tracing is skipped.

Detection results are cached by \texttt{instance\_id} so that repeated runs of the same instance do not re-scan the script.


\section{Hook Injection}
\label{sec:hook-injection}

\subsection{CPython tracing API}

CPython exposes a low-level tracing interface through \texttt{sys.settrace}. Any Python function can register a trace callback, which the interpreter then calls on three frame events: \texttt{call} (a function is entered), \texttt{line} (a source line is about to execute), and \texttt{return} (a function is about to return). The callback receives the live frame object, giving access to the current file, line number, function name, and local variable dictionary. Because the callback runs inside the interpreter loop, it observes every statement without any modification to the code under test.

\subsection{Pytest injection via conftest.py}

In a pytest project, \texttt{conftest.py} is a plugin file that pytest discovers and loads automatically before running tests in its directory. Placing hooks in the root \texttt{conftest.py} therefore affects every test in the project without touching the test files themselves.

We inject tracing by copying our tracer script into the project root as \texttt{conftest.py}. If the project already has a \texttt{conftest.py}, we back it up first. Our tracer defines two pytest hooks. The first hook, \texttt{pytest\_runtest\_call}, wraps each test execution: it installs our trace function via \texttt{sys.settrace} before the test runs and restores the previous one afterwards. The second hook, \texttt{pytest\_runtest\_makereport}, fires after each test and writes the collected data to an in-memory store when the test has failed. At session end, the \texttt{pytest\_sessionfinish} hook writes the entire store to \texttt{auto\_debug.json}.

\subsection{Other frameworks}

For unittest and Django projects, conftest.py is not available. We use a different injection mechanism that exploits CPython's module initialisation order. Before the tests run, the harness prepends \texttt{/opt/tracers} to \texttt{PYTHONPATH} and exports \texttt{AUTO\_DEBUG\_FRAMEWORK} with the detected framework name. CPython imports \texttt{sitecustomize.py} automatically from any directory on \texttt{PYTHONPATH} during interpreter startup, before any user code runs. Our \texttt{sitecustomize.py} reads \texttt{AUTO\_DEBUG\_FRAMEWORK} and dispatches to the appropriate injector: for unittest it calls \texttt{inject\_unittest\_tracer()}, which monkey-patches \texttt{unittest.TestResult} methods (\texttt{startTest}, \texttt{stopTest}, \texttt{addError}, \texttt{addFailure}) to install and remove the \texttt{sys.settrace} callback around each test; for Django it calls the equivalent \texttt{inject\_django\_tracer()}. Because the injection happens before the test runner starts, neither the project code nor the runner needs to be modified. Both wrappers produce output in the same \texttt{auto\_debug.json} format as the pytest tracer.

\subsection{Collected events}

Regardless of framework, the trace callback records three event types:

\begin{itemize}
    \item \textbf{call} events: the tracer appends a record with the file path, function name, and entry line number. No local variables are captured at this point.
    \item \textbf{line} events: the tracer appends a full frame snapshot including local variables at the moment that source line is reached.
    \item \textbf{return} events: the tracer appends a full frame snapshot including local variables at the point of return.
\end{itemize}

These three collections correspond directly to the three data structures described in \cref{sec:data-structures}.


\section{Collected Data Structures}
\label{sec:data-structures}

At session end the tracer writes all collected data to a file named \texttt{auto\_debug.json}. This file is a JSON array where each element corresponds to one failing test. An element contains the test's node identifier, the exception type and message, and three runtime data structures.

We define two record types. A \emph{call record} captures a function entry event:
\[
  r^{\mathrm{call}} = (f,\; l,\; n)
\]
where $f$ is a file path, $l \in \mathbb{N}$ is a line number, and $n$ is a function name. A \emph{frame record} extends this with a locals snapshot:
\[
  r^{\mathrm{frame}} = (f,\; l,\; n,\; \rho)
\]
where $\rho : \mathcal{V} \rightharpoonup \Sigma^{*}$ is a partial map from variable names to their serialized string values. A \emph{trace record} for one failing test is the tuple:
\[
  T = (\mathit{id},\; \tau,\; m,\; E,\; F,\; S)
\]
where $\mathit{id}$ is the test node identifier, $\tau$ is the exception type, $m$ is the exception message, and $E$, $F$, $S$ are finite sequences:
\[
  E \in (r^{\mathrm{call}})^{*}, \qquad F \in (r^{\mathrm{frame}})^{*}, \qquad S \in (r^{\mathrm{frame}})^{*}
\]

\textbf{exec\_path} ($E$) is built from \texttt{call} events and records only $(f, l, n)$ with no locals. It gives the LLM a compact call trace of the functions invoked during the test.

\textbf{frames} ($F$) is built from \texttt{return} events and records the full $(f, l, n, \rho)$ at each function return. It provides a traceback enriched with the variable state at the moment of return.

\textbf{step\_frames} ($S$) is built from \texttt{line} events and records $(f, l, n, \rho)$ at every executed source line, in order. It exposes how local variables evolve across individual statements inside a function.

Initially, we collected only entry and exit frames, believing that knowing which functions were called and what state they returned in would be sufficient to analyse runtime behaviour. This proved inadequate in practice. Empirically we observed that models given only entry and exit information struggled to localise bugs that corrupted state midway through a function, and logically this makes sense: entry frames carry no variables at all, and exit frames show only the final state at return. A bug that overwrites a variable partway through a loop or a conditional branch is invisible to both. To give the model a more granular view of what happened inside each function, we added line-level recording. With \texttt{step\_frames}, the LLM can request the execution history of a specific function and see how each local variable evolves statement by statement.

The top-level structure of a single entry in \texttt{auto\_debug.json} is shown in Listing~\ref{lst:trace-entry}.

\begin{figure}[ht]
\begin{verbatim}
{
  "nodeid":      "tests/test_foo.py::TestFoo::test_bar",
  "exc_type":    "AssertionError",
  "message":     "expected 3, got 4",
  "frames":      [ { "file": "...", "line": 42,
                     "func": "compute", "locals": {...} } ],
  "exec_path":   [ { "file": "...", "line": 10,
                     "func": "compute" } ],
  "step_frames": [ { "file": "...", "line": 11,
                     "func": "compute", "locals": {...} } ]
}
\end{verbatim}
\caption{Structure of a single entry in \texttt{auto\_debug.json}.}
\label{lst:trace-entry}
\end{figure}


\section{Filtering Pipelines}
\label{sec:filtering}

Raw traces contain frames from the Python standard library, third-party packages, pytest internals, and the tracing infrastructure itself. These frames carry no useful debugging signal and inflate the data. We apply three separate filtering pipelines, one per data structure, because the noise profile and volume differ across them. Each pipeline is a composition of custom heuristic filters that express domain knowledge about which frames are relevant for debugging application code.

Formally, a \emph{filter} is a predicate $\varphi : R \to \{0, 1\}$ over records. A \emph{pipeline} $P = (\varphi_1, \dots, \varphi_k, N)$ maps an input sequence $\sigma$ to a filtered, capacity-capped subsequence:
\[
  P(\sigma) \;=\; \bigl[\, r \in \sigma \;\mid\; \forall i:\, \varphi_i(r) = 1 \,\bigr]_{0:N}
\]
where $[\cdot]_{0:N}$ denotes truncation to the first $N$ elements. We instantiate three pipelines differing in their filter sets and cap $N$, summarised in Table~\ref{tab:pipelines}.

\begin{table}[ht]
\centering
\caption{Filtering pipelines applied per data structure.}
\label{tab:pipelines}
\begin{tabular}{@{}llr@{}}
\toprule
\textbf{Data structure} & \textbf{Filters} ($\varphi_1, \dots, \varphi_k$) & \textbf{Cap $N$} \\
\midrule
\texttt{frames}      & FrozenOrSynthetic, SitePackages, Stdlib, Conftest   & $\infty$ \\
\texttt{exec\_path}  & TestbedOnly, Conftest, Dedup$(f, n)$, MaxEntries    & 500 \\
\texttt{step\_frames}& TestbedOnly, Conftest, Dedup$(f, n, l)$, MaxEntries & 2000 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Traceback frame pipeline}

The pipeline for \texttt{frames} applies four filters in sequence. \texttt{FrozenOrSyntheticFrameFilter} drops frames whose file path contains \texttt{<} or the substring \texttt{frozen}; these correspond to frozen importlib modules that have no real source. \texttt{SitePackagesFrameFilter} drops frames whose path contains \texttt{site-packages}. \texttt{StdlibFrameFilter} drops frames from the standard library by checking whether the path starts with a known stdlib location (obtained from \texttt{sysconfig.get\_path}) or contains \texttt{/lib/python}. \texttt{ConftestFrameFilter} drops frames from \texttt{conftest.py} files, which belong to the tracing infrastructure itself.

\subsection{Execution path pipeline}

The pipeline for \texttt{exec\_path} applies \texttt{TestbedOnlyFrameFilter} (keeps only paths from the project under test, excluding \texttt{site-packages}), \texttt{ConftestFrameFilter}, \texttt{DedupFrameFilter} keyed on \texttt{(file, func)}, and \texttt{MaxEntriesFrameFilter} with a cap of 500. The dedup step collapses recursive or repeated calls to the same function into a single entry. Because \texttt{exec\_path} records have no locals, the 500-entry cap is sufficient to represent the call graph without truncating critical paths.

\subsection{Step frames pipeline}

The pipeline for \texttt{step\_frames} applies \texttt{TestbedOnlyFrameFilter}, \texttt{ConftestFrameFilter}, \texttt{DedupFrameFilter} keyed on \texttt{(file, func, line)}, and \texttt{MaxEntriesFrameFilter} with a cap of 2000. The dedup key includes the line number so that only the first execution of each source line is retained. This preserves the initial variable state at each line and avoids accumulating thousands of identical snapshots from loops. The higher cap reflects that step frames are accessed selectively by the LLM: the \texttt{get\_granular\_frames} tool accepts a function name and returns only the step frames for that function, so the full list is rarely serialized at once.


\section{Local Variable Serialization}
\label{sec:serialization}

At collection time, each \texttt{locals} dictionary is serialized immediately. Recall that $\rho : \mathcal{V} \rightharpoonup \Sigma^{*}$ maps variable names to serialized strings. We implement it through a serialization function $\sigma : \mathrm{Val} \to \Sigma^{*}$ composed with a truncation operator $\tau_k(s) = s[0:k]$:
\[
  \rho(x) \;=\; \tau_{1000}\!\bigl(\sigma(v_x)\bigr)
\]
where $v_x$ is the runtime value bound to name $x$. This is the only representation of program state the LLM ever sees: when it calls \texttt{get\_frames} or \texttt{get\_granular\_frames}, each returned $r^{\mathrm{frame}}$ carries $\rho$ entries produced by this composition. We evaluated two choices for $\sigma$.

\medskip
$\sigma_{\mathrm{repr}}(v) = \texttt{repr}(v)$. This requires no dependencies and always produces output. However, its output is Python-specific syntax. For custom objects, \texttt{repr()} returns whatever the class author implemented, which may be a truncated string, a pointer address, or may raise an exception. Nested collections produce deeply nested Python literals with no consistent schema, so the LLM must treat them as natural language rather than structured data.

\medskip
$\sigma_{\mathrm{jp}}(v) = \texttt{jsonpickle.dumps}(v,\; \texttt{unpicklable=False})$. Setting \texttt{unpicklable=False} strips the class metadata that \texttt{jsonpickle} normally embeds and produces clean JSON objects, arrays, and primitives. This output is structured and consistent across types. If \texttt{jsonpickle} is not installed in the container, the tracer falls back to $\sigma_{\mathrm{repr}}$.

\medskip
The truncation bound $\tau_{1000}$ applies after either choice. Long lists, large data frames, or deeply nested objects can produce serialized strings tens of thousands of characters long. Truncating to 1000 characters retains the leading structure of most values while preventing any single variable from dominating the context window.


\section{Integration with LLMs}
\label{sec:llm-integration}

\subsection{Trace selection}
\label{subsec:trace-selection}

A project instance may have multiple failing tests, each producing its own entry in \texttt{auto\_debug.json}. For the LLM session we select a single representative trace. The current heuristic selects the first entry in the list. This is a deliberate simplification: more sophisticated strategies such as selecting the trace with the deepest frame stack, the most distinct exception type, or the most variable-rich locals were considered but not implemented. The first-trace heuristic is deterministic and easy to reason about, which is important for reproducibility.

\subsection{Initial static prompt approach}
\label{subsec:static-prompt}

The first iteration of LLM integration built a single prompt from the collected trace data and passed it to the model in one shot. The prompt was assembled from four XML-tagged sections: \texttt{<intro>} with a task description, \texttt{<exception>} with the exception type and message, \texttt{<runtime\_trace>} with the serialized frames, and \texttt{<instructions>} with output format requirements. For each frame in the trace, the prompt builder opened the corresponding source file on the host filesystem and included ten lines of context around the frame's line number.

This approach has three problems. First, it requires access to the project's source files on the host at prompt-generation time, which is unavailable in the containerized evaluation environment used for SWE-bench instances. Second, the prompt is fixed: once generated, the model cannot ask follow-up questions or inspect parts of the code that the prompt did not include. Third, the ten-line context windows around each frame are included regardless of relevance, which can fill the context window with code the model does not need.

\subsection{Tool-based approach}
\label{subsec:tool-based}

We replaced the static prompt with an interactive tool session. The model receives a short initial prompt describing the failing test and is given access to a set of tools it can call iteratively before submitting a patch. This mirrors how a human developer investigates a bug: they look at the error, navigate to the relevant code, inspect variable states, search for symbol definitions, and only then write a fix.

The session runs for at most \texttt{max\_tool\_turns} turns (default 10). On each turn the model issues one or more tool calls. The system executes each call, returns the result, and appends both to the message history. When the model calls \texttt{apply\_patch}, the session ends immediately and the patch is extracted from the tool argument. When the model calls \texttt{get\_frames} or \texttt{get\_granular\_frames}, each frame in the response includes surrounding source lines rendered from the pre-loaded source map alongside the local variable snapshot, so the model gets code context on demand rather than all at once.

\subsection{Available tools}
\label{subsec:tools}

Table~\ref{tab:tools} lists all nine tools. They are grouped into two catalogs. The \emph{runtime catalog} contains all nine tools and is used when trace data is available. The \emph{static catalog} contains only the four project-inspection tools (\texttt{print\_project\_tree}, \texttt{open\_file}, \texttt{search\_symbol}, \texttt{apply\_patch}) and is used for baselines or ablations where no runtime data was collected.

\begin{table}[ht]
\centering
\caption{LLM tools exposed during the tool session.}
\label{tab:tools}
\begin{tabular}{@{}lp{8.2cm}@{}}
\toprule
\textbf{Tool name} & \textbf{Description} \\
\midrule
\texttt{print\_project\_tree}  & Prints the directory tree of the project, listing folders and \texttt{.py} files. \\
\texttt{open\_file}            & Reads a range of lines from a source file given a start and end line number. \\
\texttt{search\_symbol}        & Searches all \texttt{.py} files for a literal symbol string. \\
\texttt{get\_test\_error}      & Returns the exception type and message from the selected trace. \\
\texttt{get\_execution\_trace} & Returns the serialized \texttt{exec\_path}: the list of functions called during the test. \\
\texttt{get\_frames}           & Returns the serialized \texttt{frames}: return-event snapshots with locals and source context. \\
\texttt{get\_granular\_frames} & Returns the \texttt{step\_frames} for a named function: line-level snapshots with locals. \\
\texttt{get\_variable\_change} & Intended to return variable value transitions over execution. Currently unimplemented. \\
\texttt{apply\_patch}          & Accepts a unified diff patch, terminates the tool session, and submits the fix. \\
\bottomrule
\end{tabular}
\end{table}

The \texttt{apply\_patch} tool acts as the session terminator. Once the model calls it with a valid unified diff, the runner extracts the patch and stops processing further turns. If the model exhausts all turns without calling \texttt{apply\_patch}, the system attempts to extract a diff from the final assistant message as a fallback. If no patch can be found, the session returns an empty string and the instance is counted as a failure.