\chapter{Discussion}
\label{ch:discussion}

\section{How does runtime information affect LLM-based code generation? (RQ1)}
\section{What runtime information yields the best performance? (RQ2)}
\section{Is fine-tuning worth it? (RQ3)}
\section{Threats to validity}

\paragraph{Training data contamination.}
Large language models are trained on data scraped from the internet, which almost certainly includes the GitHub repositories used in SWE-bench Lite along with their associated issues and pull requests.
A model may have seen the exact bug fix during training, making the solution a form of memorization rather than real performance.
\citep{xia2024agentless} make the same observation. However, since this threat applies equally to both of our variants (\texttt{with\_runtime}, \texttt{without\_runtime}), the relative comparison is less affected than absolute numbers.

\paragraph{Benchmark representativeness.}
SWE-bench Lite covers 11 Python repositories, and results may not generalize to other languages, or bug types absent from the benchmark.
The benchmark skews toward open-source, well-tested Python projects with clear issue descriptions, which may not reflect typical software maintenance tasks. An especially important assumption is availability of tests, and informativeness of runtime information.

\paragraph{Prompt sensitivity.}
LLM performance is sensitive to prompt wording, and different system prompts or tool descriptions might produce different results.
Our prompts (see \Cref{app:prompts}) were tuned informally, and a more exhaustive prompt search might change the numbers. Potentially, prompt-optimizing frameworks such as DSPy \citep{khattab2023dspy} could be used.

\paragraph{Evaluation metric limitations.}
The binary pass/fail criterion rewards any patch that passes the test suite, including ones that special-case inputs rather than fixing the underlying logic, and does not measure code quality or generalization.
Localization accuracy measures overlap with the gold patch, but a correct fix at a different code location would be penalized despite being valid.
