\chapter{Evaluation}
\label{ch:evaluation}

\section{Experimental Setup}
\label{sec:eval-setup}

We evaluate on SWE-bench Lite~\citep{jimenez2023swebench}, a curated subset of 300 instances drawn from 11 open-source Python repositories:

\begin{center}
\begin{tabular}{ll}
\texttt{astropy}            & \texttt{mwaskom/seaborn}   \\
\texttt{django}             & \texttt{psf/requests}      \\
\texttt{flask}              & \texttt{pydata/xarray}     \\
\texttt{matplotlib}         & \texttt{pylint-dev/pylint} \\
\texttt{pallets/flask}      & \texttt{pytest-dev/pytest} \\
\texttt{scikit-learn}       &                            \\
\end{tabular}
\end{center}

Each instance corresponds to a GitHub issue with a failing test that was introduced by a real commit. The benchmark defines a binary resolution criterion: an instance is \emph{resolved} if the submitted patch causes all originally failing tests to pass without breaking any previously passing tests.

Each instance runs inside an isolated Docker container managed by \texttt{TracedInstanceRunner}, as described in \Cref{ch:system-methodology}. We run each instance under two \emph{variants}:

\begin{itemize}\setlength{\itemsep}{6pt}
  \item \textbf{without\_runtime}: the LLM session uses only the four static tools (\texttt{print\_project\_tree}, \texttt{open\_file}, \texttt{search\_symbol}, \texttt{apply\_patch}). No tracing is performed. This serves as the baseline.
  \item \textbf{with\_runtime}: the full nine-tool catalog is available, including the runtime tools (\texttt{get\_test\_error}, \texttt{get\_execution\_trace}, \texttt{get\_frames}, \texttt{get\_granular\_frames}). The tracer runs before the LLM session and populates \texttt{auto\_debug.json}.
\end{itemize}

We evaluate both variants across the following base models:

\begin{itemize}\setlength{\itemsep}{6pt}
  \item \textbf{DeepSeek}: \texttt{deepseek-chat}, \texttt{deepseek-v4-flash}, \texttt{deepseek-v4-pro}
  \item \textbf{OpenAI}: \texttt{\xxx{gpt-4o}}, \texttt{\xxx{o1}} 
  \item \textbf{Open-source}: \texttt{[TODO: qwen, llama?]}
\end{itemize}

\change{We are going to limit the scope of models. No reason to test across models of comparable performance. But we need to add small models too, so then we can compare the added value of runtime info in big vs. small models}

All models are accessed via OpenAI-compatible APIs. The default number of tool runs in session runs is 10.

We run a single LLM session per instance per variant. We do not ensemble or run multiple trials. We set temperature to 0 where the API allows it to make runs deterministic, though not all providers guarantee identical outputs at temperature 0. We report results on the full 300-instance set unless otherwise noted.

\section{Evaluation Metrics}
\label{sec:eval-metrics}

Let $\mathcal{I}$ be the set of benchmark instances, $\mathcal{M}$ a base model, and $v \in \{\texttt{with\_runtime},\, \texttt{without\_runtime}\}$ a variant. Each run produces a status $s(i, \mathcal{M}, v) \in \{\texttt{passed}, \texttt{failed}, \texttt{apply\_failed}\}$ for each instance $i \in \mathcal{I}$.

\paragraph{Resolved rate.} The primary metric is:
\[
  R(\mathcal{M}, v) \;=\; \frac{\bigl|\{\, i \in \mathcal{I} \;\mid\; s(i, \mathcal{M}, v) = \texttt{passed} \,\}\bigr|}{|\mathcal{I}|}
\]
This matches the standard SWE-bench \% resolved metric and allows direct comparison with published results.

\paragraph{Localization accuracy.} The gold patch for each instance specifies which files, functions, and lines were modified to fix the bug. We compare the patch produced by the model against the gold patch and compute accuracy at three granularities:
\[
  L_g(\mathcal{M}, v) \;=\; \frac{|\{\, i \in \mathcal{I} \;\mid\; \mathrm{loc}_g(i, \mathcal{M}, v) = 1 \,\}|}{|\mathcal{I}|}, \quad g \in \{\texttt{file},\, \texttt{func},\, \texttt{line}\}
\]
where $\mathrm{loc}_g(i, \mathcal{M}, v) = 1$ if the model's patch touches at least one of the correct files (resp. functions, lines) identified in the gold patch. File-level localization is the weakest criterion and line-level the strictest.

\paragraph{Token count.} We report the mean input and output token counts per instance:
\[
  T_{\mathrm{in}}(\mathcal{M}, v) \;=\; \frac{1}{|\mathcal{I}|}\sum_i n_{\mathrm{in}}(i, \mathcal{M}, v), \qquad
  T_{\mathrm{out}}(\mathcal{M}, v) \;=\; \frac{1}{|\mathcal{I}|}\sum_i n_{\mathrm{out}}(i, \mathcal{M}, v)
\]
Token counts are a model-independent efficiency signal. Unlike cost, they do not depend on per-provider pricing, so they remain comparable across providers and over time as prices change. Comparing $T_{\mathrm{in}}$ across variants reveals how much context the runtime tools add to the prompt. Comparing $T_{\mathrm{out}}$ shows whether the model produces longer or shorter responses when runtime data is available.

\section{Runtime Information Impact (RQ1)}
% Main results table: with vs without runtime tools, across models and both benchmarks

\section{Runtime Data Structure Ablation (RQ2)}
% exec_path only vs frames only vs step_frames vs all combinations
% Which structures contribute most to resolved rate

\section{Localization Analysis}
% File-level, function-level, line-level accuracy against gold patch
% Correlation between localisation accuracy and resolved rate

\section{Token Usage and Efficiency}
% Avg tokens per instance across configurations
% Tool call distribution: which tools are called most often

\section{Results on BugsInPy}
% Generalisability beyond SWE-Bench Lite
% Same metrics, compare trends with SWE-Bench results

