
\newif\ifEN
\ENtrue   % uncomment this for english

\def\ThesisTitleStyle{mff} % MFF style
%\def\ThesisTitleStyle{cuni} % uncomment for old-style with cuni.cz logo
%\def\ThesisTitleStyle{natur} % uncomment for nature faculty logo

\def\UKFaculty{Faculty of Mathematics and Physics}

\def\UKName{Charles University in Prague} % this is not used in the "mff" style

\def\ThesisTypeTitle{\ifEN BACHELOR THESIS \else BAKALÁŘSKÁ PRÁCE \fi}
\def\ThesisGenitive{\ifEN bachelor \else bakalářské \fi}
\def\ThesisAccusative{\ifEN bachelor \else bakalářskou \fi}

\def\ThesisTitle{Using runtime information to improve LLM code generation}
\def\ThesisAuthor{Tymofii Shchetilin}
\def\YearSubmitted{2026}

% department assigned to the thesis
\def\Department{Department of Distributed and Dependable Systems}
\def\DeptType{Department}

\def\Supervisor{Mgr. Tomáš Petříček, Ph.D.}
\def\SupervisorsDepartment{Department of Distributed and Dependable Systems}

\def\StudyProgramme{Computer Science}
\def\StudyBranch{Artificial Intelligence}

% TODO:
\def\Dedication{%
Dedication. \xxx{It is nice to say thanks to supervisors, friends, family, book authors and food providers.}
}

% TODO:
\def\AbstractEN{%
\xxx{Abstracts are an abstract form of art. Use the most precise, shortest sentences that state what problem the thesis addresses, how it is approached, pinpoint the exact result achieved, and describe the applications and significance of the results. Highlight anything novel that was discovered or improved by the thesis. Maximum length is 200 words, but try to fit into 120. Abstracts are often used for deciding if a reviewer will be suitable for the thesis; a well-written abstract thus increases the probability of getting a reviewer who will like the thesis.}
% ABSTRACT IS NOT A COPY OF YOUR THESIS ASSIGNMENT!
}

% TODO:
\def\AbstractCS{%
\xxx{You will need to submit both Czech and English abstract to the SIS, no matter what language you use in the thesis. If writing in English, translate the contents of \texttt{\textbackslash{}AbstractEN} into this field. In case you do not speak czech, your supervisor should be able to help you with the translation.}
}

% TODO:
% 3 to 5 keywords (recommended), each enclosed in curly braces.
% Keywords are useful for indexing and searching for the theses by topic.
\def\Keywords{%
\xxx{{key} {words}}
}

\def\InfoPageFont{}
